### 2017-10-18 Draft2/Appending.Dynamic.tex
Rename some commands to PascalCase
More casing
Hard-code O and P
Renaming variables

\dynarrayimpl

I will first implement appending for dynamic arrays. Let $\listname$ be a dynamic array. The following definitions are used in the code:

\begin{description}
	\item[initial capacity] The capacity of a dynamic a array when one item is appended to it. I denote this $\initcapacity$.\\
	(The capacity of any dynamic array with size zero should be zero.)
	\item[growth factor] The factor by which the current capacity is multiplied to get the new capacity when $\listname$ grows. I denote this with $\growthfactor$.
\end{description}

\begin{algorithm}
	\begin{algorithmic}[1]
		\Procedure{$\FuncAppend$}{$L, item$}
			\If{$\listname.\FieldFull$}
				\State $\listname.\FuncGrow()$
			\EndIf

			\State $\listname.\FieldHead[\listname.\FieldHeadSize] \gets item$
			\State $\listname.\FieldSize \gets \listname.\FieldSize + 1$
		\EndProcedure
		\Statex
		\Procedure{$\FuncGrow$}{$L$}
			\If{$\listname.\FieldEmpty$}
				\State $\listname.\FieldBuffer \gets \FuncNewArray(\initcapacity)$
			\Else
				\State $new\ buf \gets \FuncNewArray(\listname.\FieldSize \cdot \growthfactor)$
				\State $\FuncArrayCopy(source: \listname.\FieldBuffer, dest: new\ buf, len: \listname.\FieldSize)$
				\State $\listname.\FieldBuffer \gets new\ buf$
			\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

\tcomplex

@@

[lemma]
Let $\listname$ by a dynamic array. As $\varnitems$ are appended, the sequence of capacities $\listname$ takes on is $$S_C = 0,\ \initcapacity,\ \growthfactor\initcapacity,\ \growthfactor^2\initcapacity,\ \ldots\ \growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil}\initcapacity$$.

[proof]
The capacity starts out at zero and becomes $\initcapacity$ once one element is appended. From then on, it can only grow by a factor of $\growthfactor$ via $\FuncGrow$, so if $\growthfactor^i\initcapacity$ is the current capacity then $\growthfactor^{i + 1}\initcapacity$ must be the next capacity. By induction, the rest of the sequence is $$\left\{ \growthfactor^i\initcapacity \right\}_{i = 1}^k$$ for some $k$. Since the capacity is only as big as is needs to be, namely greater than or equal to $\varnitems$, $k$ is the smallest integer for which $\growthfactor^k\initcapacity \geq \varnitems$. Using this property, I now derive $k$.

\begin{align*}
\growthfactor^k\initcapacity &\geq \varnitems\\
\growthfactor^k &\geq \frac{\varnitems}{\initcapacity}\\
k &\geq \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity\\
k &\geq \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity > k - 1\\
k &= \big\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \big\rceil
\end{align*}

Then the final term in the sequence is $\growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil}\initcapacity$, completing the proof.

[make lemma]
Now, consider that to grow by $\growthfactor$, $\FuncGrow$ must be called. So $\growthfactor^k\initcapacity$ being present in $S_C$ means $\FuncGrow$ was called at size $\growthfactor^{k - 1}\initcapacity$ for $k \geq 1$. Then $\FuncGrow$ was called at the sizes $$S_R = 0,\ \initcapacity,\ \growthfactor\initcapacity,\ \growthfactor^2\initcapacity,\ \ldots\ \growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil - 1}\initcapacity$$

@@

\begin{align*}
0 + \initcapacity + \growthfactor\initcapacity + \ldots + \growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil - 1}\initcapacity &= \left( \frac{\growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil} - 1}{\growthfactor - 1} \right) \initcapacity
\end{align*}

@@

[ref1]
$$
\nwritesfn(\varnitems) = \varnitems + \left( \frac{\growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil} - 1}{\growthfactor - 1} \right) \initcapacity
$$

[ref2]
To make this expression easier to work with, I will write it in big-P notation to approximate it for large $\varnitems$. First I note that $\log_{\growthfactor} k \leq \lceil \log_{\growthfactor} k \rceil < \left( \log_{\growthfactor} k \right) + 1$ for any $k > 0$, so $k \leq \growthfactor^{\lceil \log_{\growthfactor} k \rceil} < \growthfactor{k}$. Then

\begin{align*}
\nwritesfn(\varnitems) &= \varnitems + \left( \frac{\growthfactor^{\big\lceil \log_{\growthfactor} \left( \frac {\varnitems} {\initcapacity} \right) \big\rceil} - 1}{\growthfactor - 1} \right) \initcapacity\\
\varnitems + \left( \frac{\frac{\varnitems}{\initcapacity} - 1}{\growthfactor - 1} \right) \initcapacity \leq \nwritesfn(\varnitems) &< \varnitems + \left( \frac{\growthfactor\frac{\varnitems}{\initcapacity} - 1}{\growthfactor - 1} \right) \initcapacity\\
\varnitems + \left( \frac{\varnitems - \initcapacity}{\growthfactor - 1} \right) \leq \nwritesfn(\varnitems) &< \varnitems + \left( \frac{\growthfactor\varnitems - \initcapacity}{\growthfactor - 1} \right)\\
\biggo\left( \varnitems + \left( \frac{\varnitems - \initcapacity}{\growthfactor - 1} \right) \right) \leq \biggo\left( \nwritesfn(\varnitems) \right) &< \biggo\left( \varnitems + \left( \frac{\growthfactor\varnitems - \initcapacity}{\growthfactor - 1} \right) \right)\\
\biggo\left( \left( \frac{\growthfactor}{\growthfactor - 1} \right) \varnitems \right) \leq \biggo\left( \nwritesfn(\varnitems) \right) &< \biggo\left( \left( \frac{2\growthfactor - 1}{\growthfactor - 1} \right) \varnitems \right)
\end{align*}

\scomplex

@@

From [lemma] I know the exact capacities $\listname$ takes on as $\varnitems$ items are appended. Dynamic arrays also have the property that $i \in S_C$ iff an array of length $i$ was allocated. Then the total length of arrays allocated is

\begin{align*}
\spacefn(\varnitems) &= \sum_{i \in S_C} i\\
&= 0 + \initcapacity + \growthfactor\initcapacity + \growthfactor^2\initcapacity + \ldots + \growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil}\initcapacity\\
&= \left( \frac{\growthfactor^{\lceil \log_{\growthfactor} \varnitems - \log_{\growthfactor} \initcapacity \rceil + 1} - 1}{\growthfactor - 1} \right) \initcapacity
\end{align*}

@@

\begin{align*}
\left( \frac{\growthfactor \frac{\varnitems}{\initcapacity} - 1}{\growthfactor - 1} \right) \initcapacity \leq \spacefn(\varnitems) &< \left( \frac{\growthfactor^2 \frac{\varnitems}{\initcapacity} - 1}{\growthfactor - 1} \right) \initcapacity\\
\biggo\left( \left( \frac{\growthfactor \frac{\varnitems}{\initcapacity} - 1}{\growthfactor - 1} \right) \initcapacity \right) \leq \biggo\left( \spacefn(\varnitems) \right) &< \biggo\left( \left( \frac{\growthfactor^2 \frac{\varnitems}{\initcapacity} - 1}{\growthfactor - 1} \right) \initcapacity \right)\\
\biggo\left( \left( \frac{\growthfactor}{\growthfactor - 1} \right) \varnitems \right) \leq \biggo\left( \spacefn(\varnitems) \right) &< \biggo\left( \left( \frac{\growthfactor^2}{\growthfactor - 1} \right) \varnitems \right)
\end{align*}

>>>>

\HdrDynArrayImpl

I will first implement appending for dynamic arrays. Let $\ListName$ be a dynamic array. The following definitions are used in the code:

\begin{description}
	\item[initial capacity] The capacity of a dynamic a array when one item is appended to it. I denote this $\InitCapacity$.\\
	(The capacity of any dynamic array with size zero should be zero.)
	\item[growth factor] The factor by which the current capacity is multiplied to get the new capacity when $\ListName$ grows. I denote this with $\GrowthFactor$.
\end{description}

\begin{algorithm}
	\begin{algorithmic}[1]
		\Procedure{$\FuncAppend$}{$L, item$}
			\If{$\ListName.\FieldFull$}
				\State $\ListName.\FuncGrow()$
			\EndIf

			\State $\ListName.\FieldHead[\ListName.\FieldHeadSize] \gets item$
			\State $\ListName.\FieldSize \gets \ListName.\FieldSize + 1$
		\EndProcedure
		\Statex
		\Procedure{$\FuncGrow$}{$L$}
			\If{$\ListName.\FieldEmpty$}
				\State $\ListName.\FieldBuffer \gets \FuncNewArray(\InitCapacity)$
			\Else
				\State $new\ buf \gets \FuncNewArray(\ListName.\FieldSize \cdot \GrowthFactor)$
				\State $\FuncArrayCopy(source: \ListName.\FieldBuffer, dest: new\ buf, len: \ListName.\FieldSize)$
				\State $\ListName.\FieldBuffer \gets new\ buf$
			\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

\HdrTimeComplex

@@

\begin{lemma}
\label{CapacitySequence}
	Let $\ListName$ by a dynamic array. As $\varnitems$ items are appended, the sequence of capacities $\ListName$ takes on is $$S_C = 0,\ \InitCapacity,\ \GrowthFactor\InitCapacity,\ \GrowthFactor^2\InitCapacity,\ \ldots\ \GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity$$.
\end{lemma}

\begin{proof}
	The capacity starts out at zero and becomes $\InitCapacity$ once one element is appended. From then on, it can only grow by a factor of $\GrowthFactor$ via $\FuncGrow$, so if $\GrowthFactor^i\InitCapacity$ is the current capacity then $\GrowthFactor^{i + 1}\InitCapacity$ must be the next capacity. By induction, the rest of the sequence is $$\left\{ \GrowthFactor^i\InitCapacity \right\}_{i = 1}^k$$ for some $k$. Since the capacity is only as big as is needs to be, namely greater than or equal to $\varnitems$, $k$ is the smallest integer for which $\GrowthFactor^k\InitCapacity \geq \varnitems$. Using this property, I now derive $k$.

	\begin{align*}
	\GrowthFactor^k\InitCapacity &\geq \varnitems\\
	\GrowthFactor^k &\geq \frac{\varnitems}{\InitCapacity}\\
	k &\geq \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity\\
	k &\geq \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity > k - 1\\
	k &= \big\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \big\rceil
	\end{align*}

	Then the final term in the sequence is $\GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity$, completing the proof.
\end{proof}

[make lemma]
Now, consider that to grow by $\GrowthFactor$, $\FuncGrow$ must be called. So $\GrowthFactor^k\InitCapacity$ being present in $S_C$ means $\FuncGrow$ was called at size $\GrowthFactor^{k - 1}\InitCapacity$ for $k \geq 1$. Then $\FuncGrow$ was called at the sizes $$S_R = 0,\ \InitCapacity,\ \GrowthFactor\InitCapacity,\ \GrowthFactor^2\InitCapacity,\ \ldots\ \GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil - 1}\InitCapacity$$

@@

\begin{align*}
0 + \InitCapacity + \GrowthFactor\InitCapacity + \ldots + \GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil - 1}\InitCapacity &= \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
\end{align*}

@@

[ref1]
$$
\nwritesfn(\varnitems) = \varnitems + \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
$$

[ref2]
To make this expression easier to work with, I will write it in big-P notation to approximate it for large $\varnitems$. First I note that $\log_{\GrowthFactor} k \leq \lceil \log_{\GrowthFactor} k \rceil < \left( \log_{\GrowthFactor} k \right) + 1$ for any $k > 0$, so $k \leq \GrowthFactor^{\lceil \log_{\GrowthFactor} k \rceil} < \GrowthFactor{k}$. Then

\begin{align*}
\nwritesfn(\varnitems) &= \varnitems + \left( \frac{\GrowthFactor^{\big\lceil \log_{\GrowthFactor} \left( \frac {\varnitems} {\InitCapacity} \right) \big\rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
\varnitems + \left( \frac{\frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \leq \nwritesfn(\varnitems) &< \varnitems + \left( \frac{\GrowthFactor\frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
\varnitems + \left( \frac{\varnitems - \InitCapacity}{\GrowthFactor - 1} \right) \leq \nwritesfn(\varnitems) &< \varnitems + \left( \frac{\GrowthFactor\varnitems - \InitCapacity}{\GrowthFactor - 1} \right)\\
P\left( \varnitems + \left( \frac{\varnitems - \InitCapacity}{\GrowthFactor - 1} \right) \right) \leq P\left( \nwritesfn(\varnitems) \right) &< P\left( \varnitems + \left( \frac{\GrowthFactor\varnitems - \InitCapacity}{\GrowthFactor - 1} \right) \right)\\
P\left( \left( \frac{\GrowthFactor}{\GrowthFactor - 1} \right) \varnitems \right) \leq P\left( \nwritesfn(\varnitems) \right) &< P\left( \left( \frac{2\GrowthFactor - 1}{\GrowthFactor - 1} \right) \varnitems \right)
\end{align*}

\HdrSpaceComplex

@@

From \ref{CapacitySequence} I know the exact capacities $\ListName$ takes on as $\varnitems$ items are appended. Dynamic arrays also have the property that $i \in S_C$ iff an array of length $i$ was allocated. Then the total length of arrays allocated is

\begin{align*}
\spacefn(\varnitems) &= \sum_{i \in S_C} i\\
&= 0 + \InitCapacity + \GrowthFactor\InitCapacity + \GrowthFactor^2\InitCapacity + \ldots + \GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity\\
&= \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil + 1} - 1}{\GrowthFactor - 1} \right) \InitCapacity
\end{align*}

@@

\begin{align*}
\left( \frac{\GrowthFactor \frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \leq \spacefn(\varnitems) &< \left( \frac{\GrowthFactor^2 \frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
P\left( \left( \frac{\GrowthFactor \frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \right) \leq P\left( \spacefn(\varnitems) \right) &< P\left( \left( \frac{\GrowthFactor^2 \frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \right)\\
P\left( \left( \frac{\GrowthFactor}{\GrowthFactor - 1} \right) \varnitems \right) \leq P\left( \spacefn(\varnitems) \right) &< P\left( \left( \frac{\GrowthFactor^2}{\GrowthFactor - 1} \right) \varnitems \right)
\end{align*}

### 2017-10-18 Draft2/Appending.Growth.tex
Renaming variables

I again start off by finding the write cost when $\varnitems$ are appended. I give the write cost function a slightly different name, $\nwritesnewfn(\varnitems)$, so I can compare it to $\nwritesfn(\varnitems)$ later. I also use $\growthfactornew$ and $\initcapacitynew$ instead of $\growthfactor$ and $\initcapacity$ for the growth factor and initial capacity, respectively.

@@

\begin{lemma}
	The capacity of a growth array grows by a constant factor $\growthfactornew$ after the first time it is called.
\end{lemma}

@@

	For the base case, consider the second time $\FuncGrow$ is called: the current capacity is $\initcapacitynew$, and the next (total) capacity should be $\growthfactornew\initcapacitynew$. Then a buffer of size $(\growthfactornew - 1)\initcapacitynew$ should be appended, and that is what the algorithm does.

	For $i \geq 2$, given the algorithm behaves correctly the $i$th time it is called, I wish to show it behaves correctly the $(i + 1)$th time. From the inductive hypothesis, when the algorithm is called the $(i + 1)$th time, the current capacity is $\growthfactornew^i\initcapacitynew$ and the head capacity is $(\growthfactornew^i - \growthfactornew^{i - 1})\initcapacitynew$. The next (total) capacity should be $\growthfactornew^{i + 1}\initcapacitynew$, and the next head capacity should be $(\growthfactornew^{i + 1} - \growthfactornew^i)\initcapacitynew$, which is $\growthfactornew$ times the current head capacity. The algorithm determines the capacity of the next buffer in this manner, proving the inductive step and completing the induction.
\end{proof}

@@

$$
S_R = 0,\ \initcapacitynew,\ \growthfactornew\initcapacitynew,\ \growthfactornew^2\initcapacitynew,\ \ldots\ \growthfactornew^{\lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil - 1}\initcapacitynew
$$

Let $k$ be the number of times the tail is appended to. Since the tail is a dynamic array, the number of writes it makes is $\nwritesfn(k)$. Since $\FuncGrow$ is called $|S_R|$ times and it appends to the tail each time except for the first, $k = |S_R| - 1 = \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil$. Then the formula for $\nwritesgrowfn(n)$ is

$$
\nwritesgrowfn(n) = k + \left( \frac{\growthfactor^{\lceil \log_{\growthfactor} k - \log_{\growthfactor} \initcapacity \rceil} - 1}{\growthfactor - 1} \right) \initcapacity
$$

where $\initcapacity$ $\growthfactor$ are the capacity of the \textit{tail} (not the growth array). I finally conclude that

$$
\nwritesnewfn(n) = n + k + \left( \frac{\growthfactor^{\lceil \log_{\growthfactor} k - \log_{\growthfactor} \initcapacity \rceil} - 1}{\growthfactor - 1} \right) \initcapacity
$$

@@

\begin{align*}
k &= \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil\\
O(k) &= O(\log \varnitems)
\end{align*}

@@

I determined in [...] that if $\varnitems$ are appended then the tail is appended to $k$ times, where $k = \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil$. Since the tail is a dynamic array, it allocates $\spacefn(k)$ space, where $\spacefn$ is the space function of the tail. Through a similar method to [...] it can be shown $O(\spacefn(k)) = O(k) = O(\log \varnitems)$. Then % TODO: P-analysis should come after the explicit formula.

>>>>

I again start off by finding the write cost when $\varnitems$ are appended. I give the write cost function a slightly different name, $\nwritesnewfn(\varnitems)$, so I can compare it to $\nwritesfn(\varnitems)$ later. I also use $\GrowthFactornew$ and $\InitCapacitynew$ instead of $\GrowthFactor$ and $\InitCapacity$ for the growth factor and initial capacity, respectively.

@@

\begin{lemma}
	The capacity of a growth array grows by a constant factor $\GrowthFactornew$ after the first time it is called.
\end{lemma}

@@

	For the base case, consider the second time $\FuncGrow$ is called: the current capacity is $\InitCapacitynew$, and the next (total) capacity should be $\GrowthFactornew\InitCapacitynew$. Then a buffer of size $(\GrowthFactornew - 1)\InitCapacitynew$ should be appended, and that is what the algorithm does.

	For $i \geq 2$, given the algorithm behaves correctly the $i$th time it is called, I wish to show it behaves correctly the $(i + 1)$th time. From the inductive hypothesis, when the algorithm is called the $(i + 1)$th time, the current capacity is $\GrowthFactornew^i\InitCapacitynew$ and the head capacity is $(\GrowthFactornew^i - \GrowthFactornew^{i - 1})\InitCapacitynew$. The next (total) capacity should be $\GrowthFactornew^{i + 1}\InitCapacitynew$, and the next head capacity should be $(\GrowthFactornew^{i + 1} - \GrowthFactornew^i)\InitCapacitynew$, which is $\GrowthFactornew$ times the current head capacity. The algorithm determines the capacity of the next buffer in this manner, proving the inductive step and completing the induction.
\end{proof}

@@

$$
S_R = 0,\ \InitCapacitynew,\ \GrowthFactornew\InitCapacitynew,\ \GrowthFactornew^2\InitCapacitynew,\ \ldots\ \GrowthFactornew^{\lceil \log_{\GrowthFactornew} \varnitems - \log_{\GrowthFactornew} \InitCapacitynew \rceil - 1}\InitCapacitynew
$$

Let $k$ be the number of times the tail is appended to. Since the tail is a dynamic array, the number of writes it makes is $\nwritesfn(k)$. Since $\FuncGrow$ is called $|S_R|$ times and it appends to the tail each time except for the first, $k = |S_R| - 1 = \lceil \log_{\GrowthFactornew} \varnitems - \log_{\GrowthFactornew} \InitCapacitynew \rceil$. Then the formula for $\nwritesgrowfn(n)$ is

$$
\nwritesgrowfn(n) = k + \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} k - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
$$

where $\InitCapacity$ $\GrowthFactor$ are the capacity of the \textit{tail} (not the growth array). I finally conclude that

$$
\nwritesnewfn(n) = n + k + \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} k - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
$$

@@

\begin{align*}
k &= \lceil \log_{\GrowthFactornew} \varnitems - \log_{\GrowthFactornew} \InitCapacitynew \rceil\\
O(k) &= O(\log \varnitems)
\end{align*}

@@

I determined in [...] that if $\varnitems$ are appended then the tail is appended to $k$ times, where $k = \lceil \log_{\GrowthFactornew} \varnitems - \log_{\GrowthFactornew} \InitCapacitynew \rceil$. Since the tail is a dynamic array, it allocates $\spacefn(k)$ space, where $\spacefn$ is the space function of the tail. Through a similar method to [...] it can be shown $O(\spacefn(k)) = O(k) = O(\log \varnitems)$. Then % TODO: P-analysis should come after the explicit formula.

### 2017-10-18 Draft2/FieldsProperties.tex
Renaming variables

I define some \textbf{fields} and \textbf{properties} (constant time methods that do not change state) for both types of arrays, so I may use them later to implement non-trivial methods.

>>>>

In subsequent sections, I will implement algorithms for both dynamic and growth arrays. In this section, I define fields and properties for these data structures which the algorithms will use. \textbf{Fields} are variables associated with an object that may be read from or written to. \textbf{Properties} are trivial, constant-time methods that do not change state.

If $\ListName$ is a dynamic array, then it is assumed to have the following fields:

\begin{itemize}
	\item $\ListName.\FieldBuffer$ - The \textbf{buffer}, or raw array, that $\ListName$ stores its items in.
	\item $\ListName.\FieldSize$ - The number of items in $\ListName$.
\end{itemize}

As a dynamic array, $\ListName$ is also given the following properties. $:=$ denotes a definition, as opposed to $=$ which checks equivalence. Functions that return boolean values are suffixed with ?.

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns the capacity of $\ListName$}
		\State $\ListName.\FieldCapacity := \ListName.\FieldBuffer.\FieldLength$
		\State
		\LineComment{Returns whether $\ListName$ is full}
		\State $\ListName.\FieldFull := \ListName.\FieldSize = \ListName.\FieldCapacity$
	\end{algorithmic}
\end{algorithm}

When a dynamic array is instantiated, the following code should run:

\begin{algorithm}
	\begin{algorithmic}
		\State $\ListName.\FieldBuffer \gets \FuncNewArray(0)$
		\State $\ListName.\FieldSize \gets 0$
	\end{algorithmic}
\end{algorithm}

If $\ListName$ is a growth array, then it is assumed to have the following fields:

\begin{itemize}
	\item $\ListName.\FieldHead$ - The \textbf{head} of $\ListName$. It returns the buffer we are currently adding items to.
	\item $\ListName.\FieldTail$ - The \textbf{tail} of $\ListName$. It returns a dynamic array of buffers that have already been filled. This can be thought of as a two-dimensional array.
	\item $\ListName.\FieldSize$ - The number of items in $\ListName$.
	\item $\ListName.\FieldCapacity$ - The \textbf{capacity} of $\ListName$. It returns the maximum number of items $\ListName$ can hold without resizing.
\end{itemize}

As a growth array, $\ListName$ is also given the following properties:

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns whether $\ListName$ is empty}
		\State $\ListName.\FieldEmpty := \ListName.\FieldSize = 0$
		\State
		\LineComment{Returns whether $\ListName$ is full}
		\State $\ListName.\FieldFull := \ListName.\FieldSize = \ListName.\FieldCapacity$
		\State
		\LineComment{Returns the capacity of $\FieldHead$}
		\State $\ListName.\FieldHeadCapacity := \ListName.\FieldHead.\FieldLength$
		\State
		\LineComment{Returns the size of $\FieldHead$}
		\LineComment{\textbf{Rationale:} $\FieldCapacity - \FieldHeadCapacity$ is the total capacity of the buffers in $\FieldTail$.}
		\LineComment{Then, $\FieldSize - \left( \FieldCapacity - \FieldHeadCapacity \right)$ is the number of items that were added}
		\LineComment{after depleting the buffers in $\FieldTail$.}
		\State $\ListName.\FieldHeadSize := \ListName.\FieldSize - \left(\ListName.\FieldCapacity - \ListName.\FieldHeadCapacity\right)$
	\end{algorithmic}
\end{algorithm}

When a growth array is instantiated, the following code should run:

\begin{algorithm}
	\begin{algorithmic}
		\State $\ListName.\FieldHead \gets \FuncNewArray(0)$
		\State $\ListName.\FieldTail \gets \FuncNewDynamicArray()$
		\State $\ListName.\FieldSize \gets 0$
		\State $\ListName.\FieldCapacity \gets 0$
	\end{algorithmic}
\end{algorithm}
